Capon Beamforming Applied on Moving Objects - An Analysis of Local Lateral Shift-Invariance

### 2013-12-08 HRUIofMovingObjects/jrnl_article.tex
Moving objects: done some writing.

\author{
   Jon Petter \AA{}sen, \IEEEmembership{Student Member,~IEEE,}, Andreas Austeng, \IEEEmembership{Member,~IEEE,} and Sverre Holm, \IEEEmembership{Senior Member,~IEEE}% <-this % stops a space

@@

An important property of an ultrasound imaging system is it's local spatial shift-invariance. Local spatial shift-invariance means that an object's appearance should be preserved when the object, or the ultrasound probe, is moved slightly. This invariance to local movements is what makes ultrasound imaging feasible in general, but it is specially important for quantitative techniques that correlate time-varying features, like pulsed doppler and flow imaging, and speckle tracking of tissue and blood. In this paper we analyze the lateral shift-invariance property of the Capon beamformer by studying the beamformer output when applied on simulated and \textit{in vitro} images of moving objects.

As mentioned, Capon beamforming has been shown to provide improved lateral resolution and edge definitions in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. The method accomplishes this by adjusting the array weights based on element-to-element covariance. Hence, the method yields data dependent weights. Capon beamforming is however also known for its suboptimal performance if the sample size available for covariance estimation is limited \cite{Mestre2006}, if correlated signals are present \cite{Widrow1982}, and if the assumed steering vector does not match the signal propagation vector \cite{Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying steering vector uncertainty sets (SVUS) \cite{Lorenz2005, Rubsamen2013}. The first two, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics for high resolution settings \cite{Synnevag2007a}. SVUS helps to reduce the signal-nulling effect of the Capon beamformer which appears if the signal is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the method obtains it's super resolution capabilities, but the signal can also get lost in the process. In this paper we will make use of steering vector set to reduce this self-nulling effect.

In \cite{Cox1973}, Cox derives formulas for the resolution of the Capon beamformer, and presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. Cox states the following important fact about the Capon beamformer (the $\vec{k}_3$ processor): 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to two times the two-way system resolution ($1/64$ radians) using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the scanned output when the object is located at plus and minus half the system resolution. No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) has also been added to the Capon beamformer output. A diagonal loading of $1/100$ corresponds to an signal-to-noise ratio of 20 dB before array gain is applied. The figure shows that a DAS beamformer, with a beamspacing equal to the system resolution, could yield a $-1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However if the Capon beamformer is used with the same beam density the scalloping loss could be as large as $-27$ dB. A $-1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam density is increased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1_2way.eps}
}
\caption{Theoretical drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming ($M=64$, $L=32$, $d=0.01$) applied in a farfield-narrowband setting. The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the system resolution, and for the latter the object is moved an angular distance corresponding to the system resolution.}
\label{fig:das_capon_beams}
\end{figure}

@@

In this paper we investigate the issue further, and we show that when a moving point scatterer is imaged with the Capon beamformer with the same beam spacing sufficient for DAS the predicted scalloping loss is observed. The question is then, what can we do in order to improve on the situation without affecting the imaging frame rate and processing throughout. In this paper we therefore study the scalloping loss of the Capon beamforming and investigate different methods aimed at reducing it. We also discuss how the different methods will affect the imaging frame rate and the processing complexity.

In the next section we give an introduction to Capon beamforming for medical ultrasound imaging together with a short introduction to lateral sampling and spatial shift invariance. In Section \ref{sec:capon_LLSI} the local lateral shift-invariance of the Capon beamformer is investigated, before different methods for improved local lateral shift-invariance are presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res} we present the results of applying the most promising method on \textit{in-vitro} data. Finally we discuss our findings in and Section \ref{sec:dis} and give our conclusions in Section \ref{sec:con}.

@@

\subsection{Capon Beamforming}
The Capon beamformer produce a weight vector, like the one in (\ref{eq:das}), based on the impinging signal and an optimization criteria. The minimization problem is as follow \cite{Capon1969}:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where $\mat{\hat{R}}$ is a sample covariance matrix. Hence, the resulting weight vector will minimizes the output power while maintaining unit gain in the steering direction $\vec{a}$. %The covariance matrix is usually unknown and has to be estimated from the element data. The covariance matrix in (\ref{eq:capon_optimization_criteria}) is therefore substituted with the sample covariance matrix.

@@

The lateral sampling required in an ultrasound image for a given center frequency, $f_c$, is given by the combined aperture size on transmit ($D_{tx}$) and receive ($D_{rx}$), and the wavelength of the center frequency, $\lambda_c = c/f_c$, where $c$ is the speed of sound in the body. The required Nyquist beam spacing is then \cite{Hergum2007}:
\begin{align}
\Delta < \frac{\lambda_c}{D_{tx} + D_{rx}}\frac{1}{p}, \label{eq:resolution}
\end{align}
where the over sampling factor $p$ can be used e.g. to reduce folding of frequencies $>f_c$. A $p$ factor of one has been used throughout this paper.

Equation (\ref{eq:resolution}) was used to calculate the object position or beam spacing in Fig.\,\ref{fig:das_capon_beams}. Therefore, if (\ref{eq:resolution}) is not fulfilled the scalloping loss could become visible in the image and the system will no longer be locally laterally shift-invariant. In an ultrasound image with 256 gray levels and 50 dB dynamic range, a -1 dB loss corresponds to 5 gray levels. This is approximately equal to the visibility threshold (Weber fraction of 2\%) for a grayscale image, and a loss larger than 1 dB could therefore end up being visible to the observer. In our further analysis a system would, based on this reasoning, be regarded as locally laterally shift-invariant if the loss in between beams is less than or close to -1 dB. %However, depending on viewing light conditions and e.g. the selected gray level mapping, a larger loss might be acceptable.

@@

To further quantify the shift-invariance we have calculated the mean absolute error (MAE) of the maximum point position in each beam profile versus a straight line ($f(x)=x$). We will refer to this measure as P-MAE. The MAE of the maximum point's amplitude versus the data maximum is also measured. We refer to this measure as A-MAE. The first measure quantifies the amount of geometric distortion \cite{Hergum2007}, and the latter quantifies the peak gain variation or in other words the scalloping loss. The maximum amplitude deviation, or maximum scalloping loss, along this point trace is also calculated. We present all these numbers together with the LSV-plot.

A given imaging system will then be locally laterally shift-invariant if the contours are diagonal and if the amplitude deviation among the maximum points is less than or close to -1 dB. Then there is minimal geometric distortion (low P-MAE) and the scalloping loss will hopefully not be visible for the observer (low A-MAE). %Since both effects can occur at the same time it is important to have a measure not effected by geometric distortion.

@@

In Fig.\,\ref{fig:das_shiftvariant} the beam density is reduced by a factor of two. We see how the geometric distortion and scalloping loss is increased to a level where the contours no longer are diagonal and constant. Hence, the imaging system is no longer locally laterally shift-invariant. We observe that the maximum peak is both geometrically distorted and attenuated when the point is located between two beams and the distance between the beams are too high.

@@

\section{Local Lateral Shift-Invariance of the Capon Beamformer}\label{sec:capon_LLSI}
We will now investigate the local lateral shift-invariance of the Capon beamformer with parameters equal to the ones used in earlier work \cite{Synnevag2009, Asen}. In Fig.\,\ref{fig:capon1} the same point scatterer response as presented in Fig.\,\ref{fig:das_shiftinvariant} has been weighted with Capon weights calculated using (\ref{eq:capon_weights}) with $L = M/2 = 48$, $K=1$ and $d=1/100$. As mentioned, the set of lateral displacements is constructed in such a way that the center beam is the only one hitting the point scatterer exactly. The signal cancellation caused by the difference between the signal propagation vector and the assumed steering vector is therefore minimal at this location. From the plot we observe that the signal cancellation is high in all the other lateral positions. As a consequence the scalloping loss is tremendous, and not far from the predicted value in Fig\,\ref{fig:das_capon_beams}. %The difference between the predicted and the simulated loss might be caused by several factors. The most important one being that Fig.\,\ref{fig:das_capon_beams} shows a single frequency whereas the RMS operator used in Fig.\,\ref{fig:capon1} average a broadband signal in range. For a slice through the center of the pulse the scalloping loss is below -30dB.

The diagonal factor $d$ and the effective adaptive aperture size $L$ can be used to control the mainlobe width of the Capon beamformer. In both extreme cases, $d\gg1$ and $L=1$, the Capon beamformer becomes either a triangle or uniform-weighted DAS beamformer respectively, and will require a beamspacing according to (\ref{eq:resolution}). In Fig.\,\ref{fig:capon2} we show how the scalloping loss is reduced when the amount of diagonal loading is increased.

@@

From Fig.\,\ref{fig:capon_oversampling_b} we see that more than ten times oversampling has to be applied to achieve a local lateral shift-invariant Capon beamformer when a moving point scatterer is imaged without noise and with the same parameters as in Fig\,\ref{fig:capon}. %$1/100$ in diagonal loading. 
Note that a diagonal loading factor of $1/100$ will introduce a deterministic white noise component in the sample covariance matrix $20$ dB lower than the signal present. Hence, the covariance has been estimated from data which appeared to have an SNR of 20 dB.

@@

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams in order to get sufficient sampling on receive, it is possible to achieve high frame rates in e.g. cardiac ultrasound imaging. However, PRB also introduces motion artifacts in the image which to some degree can be compensated for \cite{Hergum2007, Denarie2011, Denarie2013}. In \cite{Rabinovich2013} it is shown that synthetic transmit beamforming (STB) \cite{Hergum2007} works with Capon beamforming. In \cite{Asen} we also used STB in order to increase the beam density of a cardiac ultrasound recording. In the following discussion we consider corrected PRB as being equal to oversampling on transmit when it comes to image quality. The LSV-plots for PRB are therefore not included.

@@

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since the ultrasound data are pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole sector by only using phase rotation. This is also possible in broadband applications as long as the phase rotation is limited to a fraction of the wave length (See coarse-fine beamforming in \cite{Thomenius}). It is therefore possible to make a sweep between a given beam and its two nearest neighbours using only phase rotation without introducing to large errors. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

For an uniform linear array located along the x-axis the steering vector $\vec{a}_\theta$ should be calculated as 
\begin{align}
\vec{a}_\theta = 
\begin{bmatrix}
e^{-j\frac{2\pi}{\lambda_c}x_0\sin(\theta)} \\
e^{-j\frac{2\pi}{\lambda_c}x_1\sin(\theta)} \\
\vdots \\ 
e^{-j\frac{2\pi}{\lambda_c}x_{L-1}\sin(\theta)}
\end{bmatrix}
\end{align}
where $x_i$ is the element position and $\theta$ is swiped from $-\Delta/2$ to $\Delta/2$ where $\Delta$ is as defined in (\ref{eq:resolution}). Note that the steering vector is calculated for a $L$ element array centered around the origin, even though the full array has $M$ elements. It is important that this subarray has the same element pitch, $x_i - x_{i-1}$, as the full array. The steering error introduced by this steering vector for a frequency band $B$ is less than $B/2 \times \Delta/2$.

>>>>

\author{
   Jon Petter \AA{}sen, \IEEEmembership{Student Member,~IEEE,} Andreas Austeng, \IEEEmembership{Member,~IEEE,} and Sverre Holm, \IEEEmembership{Senior Member,~IEEE}% <-this % stops a space

@@

An important property of an ultrasound imaging system is its local spatial shift-invariance. Local spatial shift-invariance means that an object's appearance should be preserved when the object, or the ultrasound probe, is moved slightly. This invariance to local movements is what makes ultrasound imaging feasible in general, but it is specially important for quantitative techniques that correlate time-varying features, like pulsed doppler and flow imaging, and speckle tracking of tissue and blood. In this paper we analyze the lateral shift-invariance property of the Capon beamformer by studying the beamformer output when applied on simulated and \textit{in vitro} images of moving objects.

As mentioned, Capon beamforming has been shown to provide improved lateral resolution and edge definitions in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. The method accomplishes this by adjusting the array weights based on element-to-element covariance. Hence, the method yields data dependent weights. Capon beamforming is however also known for its suboptimal performance if e.g. the sample size available for covariance estimation is limited \cite{Mestre2006}, if correlated signals are present \cite{Widrow1982}, and if the assumed steering vector does not match the signal propagation vector \cite{Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying a steering vector uncertainty set \cite{Lorenz2005, Rubsamen2013}. Spatial smoothing and diagonal loading, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics for high resolution parameters \cite{Synnevag2007a}. A steering vector uncertainty sets helps to reduce signal-nulling which appears if the signal of interest is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the Capon beamformer obtains its super resolution capabilities, but the signal can also get lost in the process. In \cite{JianLi2003} it was shown that most methods based on steering vector uncertainty sets are equivalent to diagonal loading where the loading factor depends on the steering vector uncertainty. In this paper we will make use of a set of steering vectors to reduce this self-nulling effect by means of over sampling.

In \cite{Cox1973}, Cox derives formulas for the resolution of the Capon beamformer, and presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. Cox states the following important fact about the Capon beamformer (the $\vec{k}_3$ processor): 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to two times the two-way system resolution ($1/64$ radians) using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the scanned output when the object is located at plus and minus half the system resolution. No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) has also been added to the Capon beamformer output. A diagonal loading of $1/100$ corresponds to a signal-to-noise ratio (SNR) of 20 dB before array gain is applied. The figure shows that a DAS beamformer, with a beamspacing equal to the system resolution, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However if the Capon beamformer is used with the same beam density the scalloping loss could be as large as $27$ dB. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam density is increased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1_2way.eps}
}
\caption{Theoretical drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming ($M=64$, $L=32$, $d=0.01$) applied in a farfield-narrowband setting. The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the system resolution, and for the latter the object is moved an angular distance corresponding to the system resolution. If the beam density was reduced to one beam per system resolution the maximum scalloping loss is found at zero degrees.}
\label{fig:das_capon_beams}
\end{figure}

@@

In this paper we investigate the issue further. We show that when a moving point scatterer is imaged with the Capon beamformer with the same beam spacing sufficient for DAS, the predicted scalloping loss is observed. The question is then, what can we do in order to improve on the situation without affecting the imaging frame rate and processing throughout. We try two answer this by studying the scalloping loss of the Capon beamforming and investigate different methods aimed at reducing it. We also discuss how the different methods will affect the imaging frame rate and the processing complexity.

In the next section we give an introduction to Capon beamforming for medical ultrasound imaging together with a short introduction to lateral sampling and spatial shift-invariance. In Section \ref{sec:capon_LLSI} the local lateral shift-invariance of the Capon beamformer is investigated, before different methods for improved local lateral shift-invariance are presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res} we present the results of applying the most promising method on \textit{in-vitro} data. Finally we discuss our findings in Section \ref{sec:dis} and give our conclusions in Section \ref{sec:con}.

@@

\subsection{Capon Beamforming}
The Capon beamformer produce a weight vector, like the one in (\ref{eq:das}), based on the impinging signal and an optimization criterion. The minimization problem is as follow \cite{Capon1969}:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where $\mat{\hat{R}}$ is a sample covariance matrix. Hence, the resulting weight vector will minimizes the output power while maintaining unit gain in the steering direction $\vec{a}$. %The covariance matrix is usually unknown and has to be estimated from the element data. The covariance matrix in (\ref{eq:capon_optimization_criteria}) is therefore substituted with the sample covariance matrix.

@@

The lateral sampling required in an ultrasound image for a given center frequency, $f_c$, and bandwidth, $B$, is given by the combined aperture size on transmit ($D_{tx}$) and receive ($D_{rx}$), and the wavelength of the center frequency, $\lambda_c = c/f_c$, where $c$ is the speed of sound in the body. The required Nyquist beam spacing is then \cite{Hergum2007}:
\begin{align}
\Delta < \frac{\lambda_c}{D_{tx} + D_{rx}}\frac{1}{p}, \label{eq:resolution}
\end{align}
where the over sampling factor $p$ can be used to reduce folding of frequencies between $f_c$ and $f_c + B/2$. A $p$ factor of one has been used throughout this paper.

Equation (\ref{eq:resolution}) was used to calculate the object position or beam spacing in Fig.\,\ref{fig:das_capon_beams}. Therefore, if (\ref{eq:resolution}) is not fulfilled the scalloping loss could become visible in the image and the system will no longer be locally laterally shift-invariant. In an ultrasound image with 256 gray levels and 50 dB dynamic range, a 1 dB loss corresponds to 5 gray levels. This is approximately equal to the visibility threshold (Weber fraction of 2\%) for a grayscale image, and a loss larger than 1 dB could therefore end up being visible to the observer. In our further analysis a system would be regarded as locally laterally shift-invariant if the scalloping loss is less than 1 dB. However, what eventually matters is whether the loss is visible in the image or not. %However, depending on viewing light conditions and e.g. the selected gray level mapping, a larger loss might be acceptable.

@@

To further quantify the shift-invariance we have calculated the mean absolute error (MAE) of the maximum point position in each beam profile versus a straight line ($f(x)=x$). We will refer to this measure as P-MAE. The MAE of the maximum points' amplitude versus the data maximum is also measured. We refer to this measure as A-MAE. The first measure quantifies the amount of geometric distortion \cite{Hergum2007}, and the latter quantifies the peak gain variation or in other words the scalloping loss. The maximum amplitude deviation, or maximum scalloping loss, along this point trace is also calculated. We present all these numbers together with the LSV-plot.

A given imaging system will then be locally laterally shift-invariant if the contours are diagonal and if the amplitude deviation among the maximum points is less than 1 dB. Then there is minimal geometric distortion (low P-MAE) and the scalloping loss will hopefully not be visible for the observer (low A-MAE). %Since both effects can occur at the same time it is important to have a measure not effected by geometric distortion.

@@

In Fig.\,\ref{fig:das_shiftvariant} the beam density is reduced by a factor of two. We see how the geometric distortion and scalloping loss is increased to a level where the contours no longer are diagonal and the maximum trace is not constant. Hence, the imaging system is no longer locally laterally shift-invariant. We observe that the maximum peak is both geometrically distorted and attenuated when the point is located between two beams and the distance between the beams are too high.

@@

\section{Local Lateral Shift-Invariance of the Capon Beamformer}\label{sec:capon_LLSI}
We will now investigate the local lateral shift-invariance of the Capon beamformer with parameters equal to the ones used in earlier work \cite{Synnevag2009, Asen}. In Fig.\,\ref{fig:capon1} the same point scatterer response as presented in Fig.\,\ref{fig:das_shiftinvariant} has been weighted with Capon weights calculated using (\ref{eq:capon_weights}) with $L = M/2 = 48$, $K=1$ and $d=1/100$. As mentioned, the beam at zero degrees is the only one where the point scatterer is located in the center of the beam. The signal cancellation caused by the difference between the signal propagation vector and the assumed steering vector is therefore minimal when the point scatterer has zero lateral offset. From the plot we observe that the signal cancellation is high for all the other offsets. As a consequence the scalloping loss is tremendous, and not far from the predicted value in Fig\,\ref{fig:das_capon_beams}. %The difference between the predicted and the simulated loss might be caused by several factors. The most important one being that Fig.\,\ref{fig:das_capon_beams} shows a single frequency whereas the RMS operator used in Fig.\,\ref{fig:capon1} average a broadband signal in range. For a slice through the center of the pulse the scalloping loss is below -30dB.

The diagonal factor $d$ and the effective adaptive aperture size $L$ can be used to control the mainlobe width of the Capon beamformer. In both extreme cases, $d\gg1$ and $L=1$, the Capon beamformer becomes either a triangle or uniform-weighted DAS beamformer respectively, and a beamspacing according to (\ref{eq:resolution}) would then be sufficient. In Fig.\,\ref{fig:capon2} we show how the scalloping loss is reduced when the amount of diagonal loading is increased.

@@

From Fig.\,\ref{fig:capon_oversampling_b} we see that more than ten times oversampling has to be applied to achieve a local lateral shift-invariant Capon beamformer when a moving point scatterer is imaged without noise and with the same parameters as in Fig\,\ref{fig:capon}. %$1/100$ in diagonal loading. 
Note that a diagonal loading factor of $1/100$ will introduce a white noise component in the sample covariance matrix $20$ dB lower than the signal present. Hence, the covariance has been estimated from data which appeared to have an SNR of 20 dB.

@@

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams in order to get sufficient sampling on receive, it is possible to achieve high frame rates in e.g. cardiac ultrasound imaging. However, PRB also introduces geometrical distortion which to some degree can be compensated for by using synthetic transmit beamforming (STB) \cite{Hergum2007, Denarie2011, Denarie2013}. In \cite{Rabinovich2013} it is shown that STB works with Capon beamforming. In \cite{Asen} we also used STB in order to increase the beam density of a cardiac ultrasound recording. In the following discussion we consider corrected PRB as being equal to oversampling on transmit when it comes to image quality. The LSV-plots for PRB are therefore not included.

@@

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since the ultrasound data are pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole sector by using phase rotation. This is also possible in broadband applications as long as the phase rotation is limited to a fraction of the wave length (See coarse-fine beamforming in \cite{Thomenius} \comment{find a better citation}). It is therefore possible to make a sweep between a given beam and its two nearest neighbours using only phase rotation without introducing to large errors. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

For a uniform linear array located along the x-axis the steering vector $\vec{a}_\theta$ should be calculated as 
\begin{align}
\vec{a}_\theta = 
\begin{bmatrix}
e^{-j\frac{2\pi}{\lambda_c}x_0\sin(\theta)} \\
e^{-j\frac{2\pi}{\lambda_c}x_1\sin(\theta)} \\
\vdots \\ 
e^{-j\frac{2\pi}{\lambda_c}x_{L-1}\sin(\theta)}
\end{bmatrix}
\end{align}
where $x_i$ is the element position and $\theta$ is swiped from $-\Delta/2$ to $\Delta/2$ where $\Delta$ is as defined in (\ref{eq:resolution}). Note that the steering vector is calculated for a $L$ element array centered around the origin, even though the full array has $M$ elements. It is important that this subarray has the same element pitch, $x_i - x_{i-1}$, as the full array. The steering error introduced by this steering vector for a frequency band $B$ is less than $B/2 \times \Delta/2$.
